### 2016-01-26 inputs/discussion.tex
modifiche 12:38

\subsection{Minor Allele Frequency and data (un)balancedness}
\label{sec:maf}
GBS data pose a greater challenge than SNP-chip data to imputation algorithms, mainly as a consequence of the much larger quantity of missing genotypes they usually contain. On our data sets we found missing rates varying from 53\% to 67\%. With larger amounts of missing data the complexity of the imputation problem tends to increase. Moreover, with such a high level of missingness, even imputation errors as low as 1\% may have negative impact on successive analysis, amplifying the importance of finding the most accurate imputation method.\\
The imputation of missing SNP genotypes is a special case of the broader family of classification problems: three-class missing genotypes (AA, AB and BB) at any given SNP locus are classified based on known genotypes at all remaining data points. Classification problems are known to be harder when data are unbalanced, i.e. the classes appear at different frequencies in the datasets, with typically one over-represented class
%problem with url
(see~\cite{kotsiantis_handling_2006}
and~\cite{sun_classification_2009} for a review). In SNP genotype imputation, data balancedness is directly related to the minor allele frequency (MAF). In the classification of unbalanced observations, it is important to look not only at the total classification accuracy, but also at the per-class accuracy. The total classification accuracy may be misleading, being ``dominated'' by the majority class
%problem with url
\cite{he_learning_2009}.
Indeed, we found that for most methods, even when the total classification accuracy was very high, relatively large error rates were present on minority classes. 
The dependency of imputation results on MAF has already been acknowledged  (e.g.,~\cite{hickey_factors_2012} in maize; \cite{ma_comparison_2013} in cattle; \cite{pei_analyses_2008} in humans). Our per-class dissection of results allowed a deeper insight into the imputation process. Indeed, all imputation algorithms performed considerably better in the majority class rather than in the less frequent classes.
Some algorithms failed almost completely the imputation of the least common genotypes, while other algorithms proved to work well. In alfalfa, KNNI was easily the best imputation method in the heterozygous and minor homozygous classes, while SVDI and RFI performed only slightly better than MNI. In rice, Beagle gave the best imputation results, with accuracy close to 100\% both in the major and minor homozygous classes. MNI performed reasonably well in the alfalfa heterozygous class, with accuracy close to 0.6, contributed to its moderately good results (total accuracy around 0.8).

@@

Problem size, besides affecting the accuracy achievable by some imputation methods, is mainly relevant with reference to the required computation resources. RFI was the most demanding algorithm, with computation times growing exponentially both with the total number of marker genotypes (m SNP $\times$ n samples) and with the number of missing genotypes to be imputed. When analyzing the complete rice dataset (all chromosomes together), RFI took a maximum of about 40 days to complete imputation, and became computationally intractable (on the available platform) for the largest missing rate scenarios. This was partially mitigated by parallelization (we used 10 CPUs in our experiments). All other imputation algorithms took much shorter times to complete imputation, even in the most challenging scenarios. Putting together imputation accuracy and computation time, we found that the best performing imputation algorithms were Beagle in rice and KNNI in alfalfa.\\

>>>>

\subsection{Minor Allele Frequency and data (un)balancedness}
\label{sec:maf}
GBS data pose a greater challenge than SNP-chip data to imputation algorithms, mainly as a consequence of the much larger quantity of missing genotypes they usually contain. On our data sets we found missing rates varying from 53\% to 67\%. With larger amounts of missing data the complexity of the imputation problem tends to increase. Imputation error can have a negative impact on successive analyses (see for instance \cite{rutkoski_imputation_2013}). Contrary to SNP array data, where sporadic missing genotypes are around $2$-$5\%$, with the large amounts of missing data typical of GBS experiments, even imputation errors as low as 1\% may be detrimental, amplifying the importance of finding the most accurate imputation method.\\
The imputation of missing SNP genotypes is a special case of the broader family of classification problems: three-class missing genotypes (AA, AB and BB) at any given SNP locus are classified based on known genotypes at all remaining data points. Classification problems are known to be harder when data are unbalanced, i.e. the classes appear at different frequencies in the datasets, with typically one over-represented class
%problem with url
(see~\cite{kotsiantis_handling_2006}
and~\cite{sun_classification_2009} for a review). In SNP genotype imputation, data balancedness is directly related to the minor allele frequency (MAF). In the classification of unbalanced observations, it is important to look not only at the total classification accuracy, but also at the per-class accuracy. The total classification accuracy may be misleading, being ``dominated'' by the majority class
%problem with url
\cite{he_learning_2009}.
Indeed, we found that for most methods, even when the total classification accuracy was very high, relatively large error rates were present on minority classes. 
The dependency of imputation results on MAF has already been acknowledged  (e.g.,~\cite{hickey_factors_2012} in maize; \cite{ma_comparison_2013} in cattle; \cite{pei_analyses_2008} in humans). Our per-class dissection of results allowed a deeper insight into the imputation process. Indeed, all imputation algorithms performed considerably better in the majority class rather than in the less frequent classes.
Some algorithms failed almost completely the imputation of the least common genotypes, while other algorithms proved to work well. In alfalfa, KNNI was easily the best imputation method in the heterozygous and minor homozygous classes, while SVDI and RFI performed only slightly better than MNI. In rice, Beagle gave the best imputation results, with accuracy close to 100\% both in the major and minor homozygous classes. MNI performed reasonably well in the alfalfa heterozygous class, with accuracy close to 0.6, contributed to its moderately good results (total accuracy around 0.8).

@@

### 2016-01-26 inputs/materials_and_methods.tex
modifiche 26 mattina 10:36
modifiche 12:38

\begin{equation}
\label{eq:SVDI_regression}
Y = U\beta +\varepsilon 
\end{equation}

with Y as vector of genotypes at marker j; $\hat{U}$ matrix of k eigenvectors; $\beta$ vector of regression coefficients; $\varepsilon$ random errors terms; (iv) all eigenvectors (matrix U) and $\beta$ were used to estimate missing values at marker j. The procedure was repeated (steps (ii) to (iv)) until convergence. The final imputed data points were then discretized to the closer genotype in \{0,1,2\}. Additional details on SVDI can be found in \cite{troyanskaya_missing_2001}

@@

>>>>

\begin{equation}
\label{eq:SVDI_regression}
Y = U^*\beta +\varepsilon 
\end{equation}

with Y as vector of genotypes at marker j; $U^*$ matrix of the first k eigenvectors; $\beta$ vector of regression coefficients; $\varepsilon$ random errors terms; (iv) all eigenvectors (matrix U) and $\beta$ were used to estimate missing values at marker j. The procedure was repeated (steps (ii) to (iv)) until convergence. The final imputed data points were then discretized to the closer genotype in \{0,1,2\}. Additional details on SVDI can be found in \cite{troyanskaya_missing_2001}

@@

[ADD command line!!]
